---
title: "Hands-On Ex 3.1"
author: "Chai Zhixuan"
date: 2024-01-12
date-format: long
date-modified: "last-modified"
---

# **Getting Started**

This hands-on exercise is on **Programming Interactive Data Visualisation with R**

::: callout-note
## Zhixuan's personal learning outcomes

1.  Learn and create interactive data visualisation
2.  Link it back to work and potential project?
:::

The following R packages will be used:

-   [**ggiraph**](https://davidgohel.github.io/ggiraph/) for making 'ggplot' graphics interactive.

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables](https://datatables.net/) that create interactive table on html page.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

-   [**patchwork**](https://patchwork.data-imaginist.com/) for combining multiple ggplot2 graphs into one figure.

The code chunk below will be used to accomplish the task.

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

# Importing data

Let's use the exam data again 😂

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# **Interactive Data Visualisation - ggiraph methods**

I have read this [subsection](https://r4va.netlify.app/chap03#interactive-data-visualisation---ggiraph-methods) before proceeding.

### **Tooltip effect with *tooltip* aesthetic**

::: callout-note
First, an ggplot object will be created. Next, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) of **ggiraph** will be used to create an interactive svg object.
:::

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

I will try with English score

```{r}
p <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

### **Displaying multiple information on tooltip**

::: callout-note
The first three lines of codes in the code chunk create a new field called *tooltip*. At the same time, it populates text in ID and CLASS fields into the newly created field. Next, this newly created field is used as tooltip field as shown in the code of line 7.
:::

```{r}
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

I will try with Science data this time 😃

```{r}
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = SCIENCE)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

::: callout-tip
The student\'s ID and Class will be displayed now! 😊
:::

### **Customising Tooltip style**

[`opts_tooltip()`](https://davidgohel.github.io/ggiraph/reference/opts_tooltip.html) of **ggiraph** to customize tooltip rendering by add css declarations.

```{r}
tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

::: callout-note
Background colour of the tooltip is black and the font colour is white and bold.
:::

### **Displaying statistics on tooltip**

::: callout-note
A function will be used to compute 90% confident interval of the mean. The derived statistics are then displayed in the tooltip. This is good!
:::

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

I proceed to try with English score

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean english scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = ENGLISH, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = ENGLISH),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

### **Hover effect with *data_id* aesthetic**

::: callout-note
The second interactive feature of ggiraph, namely: `data_id`
:::

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        
```

::: callout-note
Interactivity: Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over. 😱

The default value of the hover css is *hover_css = \"fill:orange;\"*.
:::

### **Styling hover effect**

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

::: callout-note
The example above , the ccs customisation request are encoded directly.
:::

### **Combining tooltip and hover effect**

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

::: callout-note
Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over. At the same time, the tooltip will show the CLASS.
:::

### **Click effect with onclick**

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```

::: callout-note
Web document link with a data object will be displayed on the web browser upon mouse click.
:::

### **Coordinated Multiple Views with ggiraph**

```{r}
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

::: callout-note
Nice! This is good to compare two plots!
:::

# **Interactive Data Visualisation - plotly methods!**

### **Creating an interactive scatter plot: plot_ly() method**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| eval: true
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)

```

## The code

```{r}
#| eval: false

plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)
```
:::

### **Working with visual variable: plot_ly() method**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}

#| eval: true
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)
```

## The code

```{r}
#| eval: false

plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)

```
:::

### **Creating an interactive scatter plot: ggplotly() method**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| eval: true
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)

```

## The code

```{r}
#| eval: false

p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)
```
:::

### **Coordinated Multiple Views with plotly**

::: callout-note
The creation of a coordinated linked plot by using plotly involves three steps:

-   [`highlight_key()`](https://www.rdocumentation.org/packages/plotly/versions/4.9.2/topics/highlight_key) of **plotly** package is used as shared data.

-   two scatterplots will be created by using ggplot2 functions.

-   lastly, [*subplot()*](https://plotly.com/r/subplots/) of **plotly** package is used to place them next to each other side-by-side.
:::

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| eval: true

d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))

```

## The code

```{r}
#| eval: false

d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))
```
:::

## **Interactive Data Visualisation - crosstalk methods!**

### **Interactive Data Table: DT package**

😱

```{r}
DT::datatable(exam_data, class= "compact")
```

### **Linked brushing: crosstalk method**

😱

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| eval: true

d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        

```

## The code

```{r}
#| eval: false


d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        

```
:::
