---
title: "Hands-on Ex3.2"
author: "Chai Zhixuan"
date: 2024-01-20
date-format: long
date-modified: "last-modified"
---

# **Getting Started**

This hands-on exercise is on **Programming Animated Statistical Graphics with R**

::: callout-note
## Zhixuan's personal learning outcomes

1.  Learn and create animated statistical graphics
2.  Link it back to work and potential project?
:::

::: callout-important
Tips from Prof Kam: Before you start making animated graphs, you should first ask yourself: Does it makes sense to go through the effort? If you are conducting an exploratory data analysis, a animated graphic may not be worth the time investment. However, if you are giving a presentation, a few well-placed animated graphics can help an audience connect with your topic remarkably better than static counterparts.
:::

### **Loading the R packages**

I will load the following R packages:

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**gganimate**](https://gganimate.com/), an ggplot extension for creating animated statistical graphs.

-   [**gifski**](https://cran.r-project.org/web/packages/gifski/) converts video frames to GIF animations using pngquant\'s fancy features for efficient cross-frame palettes and temporal dithering. It produces animated GIFs that use thousands of colors per frame.

-   [**gapminder**](https://cran.r-project.org/web/packages/gapminder/): An excerpt of the data available at Gapminder.org. We just want to use its *country_colors* scheme.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate, tidyverse)
```

### **Importing the data**

Finally new dataset!! 😄

```{r}
col <- c("Country", "Continent")
globalPop <- read_xls("data/GlobalPopulation.xls",
                      sheet="Data") %>%
  mutate_at(col, as.factor) %>%
  mutate(Year = as.integer(Year))
```

## **Animated Data Visualisation: gganimate methods**

I have read the important points located [here](https://r4va.netlify.app/chap04#animated-data-visualisation-gganimate-methods)

### **Building a static population bubble plot**

::: callout-note
The basic ggplot2 functions are used to create a static bubble plot as shown below.
:::

```{r}
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') 
```

::: callout-note
Interesting plot. But I am trying to understand what it means 🙃
:::

### **Building the animated bubble plot**

::: callout-note
In the code chunk below,

-   [`transition_time()`](https://gganimate.com/reference/transition_time.html) of **gganimate** is used to create transition through distinct states in time (i.e. Year).

-   `ease_aes()` is used to control easing of aesthetics. The default is `linear`. Other methods are: quadratic, cubic, quartic, quintic, sine, circular, exponential, elastic, back, and bounce.
:::

```{r}
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') +
  transition_time(Year) +       
  ease_aes('linear')          
```

This is 😱

## **Animated Data Visualisation: plotly**

### **Building an animated bubble plot: `ggplotly()` method**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true

gg <- ggplot(globalPop, 
       aes(x = Old, 
           y = Young, 
           size = Population, 
           colour = Country)) +
  geom_point(aes(size = Population,
                 frame = Year),
             alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(x = '% Aged', 
       y = '% Young')

ggplotly(gg)
```

## The code chunk

```{r}
#| eval: false
gg <- ggplot(globalPop, 
       aes(x = Old, 
           y = Young, 
           size = Population, 
           colour = Country)) +
  geom_point(aes(size = Population,
                 frame = Year),
             alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(x = '% Aged', 
       y = '% Young')

ggplotly(gg)
```
:::

::: callout-important
Note that although `show.legend = FALSE` argument was used, the legend still appears on the plot. To overcome this problem, `theme(legend.position='none')` should be used as shown in the plot and code chunk below.
:::

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true


gg <- ggplot(globalPop, 
       aes(x = Old, 
           y = Young, 
           size = Population, 
           colour = Country)) +
  geom_point(aes(size = Population,
                 frame = Year),
             alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(x = '% Aged', 
       y = '% Young') + 
  theme(legend.position='none')

ggplotly(gg)
```

## The code chunk

```{r}
#| eval: false

gg <- ggplot(globalPop, 
       aes(x = Old, 
           y = Young, 
           size = Population, 
           colour = Country)) +
  geom_point(aes(size = Population,
                 frame = Year),
             alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(x = '% Aged', 
       y = '% Young') + 
  theme(legend.position='none')

ggplotly(gg)
```
:::

### **Building an animated bubble plot: `plot_ly()` method**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true

bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent,
          sizes = c(2, 100),
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          ) %>%
  layout(showlegend = FALSE)
bp
```

## The code chunk

```{r}
#| eval: false

bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent,
          sizes = c(2, 100),
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          ) %>%
  layout(showlegend = FALSE)
bp
```
:::
