---
title: "Hands-on Ex 4"
author: "Chai Zhixuan"
date: 2024-01-14
date-format: long
date-modified: "last-modified"
---

# **Getting Started**

::: callout-note
## Zhixuan's personal learning outcomes:

1.  Understand and do visualisation distribution
2.  Understand and apply visual statistical analysis
3.  Understand and apply visualizing uncertainty
4.  Understand and apply Funnel Plots for Fair Comparisons
5.  Apply and link it to take home exercise, project and work
:::

# **1 Visualising Distribution**

For the purpose of this exercise, the following R packages will be used, they are:

-   tidyverse, a family of R packages for data science process,

-   ggridges, a ggplot2 extension specially designed for plotting ridgeline plots, and

-   ggdist for visualising distribution and uncertainty.

```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)
```

*Exam_data.csv* will be used again 🤣

```{r}
exam <- read_csv("data/Exam_data.csv")
```

## **Visualising Distribution with Ridgeline Plot**

The ridgeline plot below is plotted by using `geom_density_ridges()`.

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()

```

## The code

```{r}
#| eval: false

ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```
:::

I will try with Math scores:

```{r}
ggplot(exam, 
       aes(x = MATHS, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "Math grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### **Varying fill colors along the x axis**

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()

```

## The code

```{r}
#| eval: false
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```
:::

::: callout-note
Colour seems to be nice
:::

### **Mapping the probabilities directly onto colour**

From prof, figure below is plotted by mapping the probabilities calculated by using `stat(ecdf)` which represent the empirical cumulative density function for the distribution of English score.

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true

ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()

```

## The code

```{r}
#| eval: false

ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()

```
:::

### **Ridgeline plots with quantile lines**

By using [`geom_density_ridges_gradient()`](https://wilkelab.org/ggridges/reference/geom_ridgeline_gradient.html), we can colour the ridgeline plot by quantile, via the calculated `stat(quantile)` aesthetic as shown in the figure below.

::: panel-tabset
## The plot

```{r, echo=FALSE, warning=FALSE}
#| eval: true


ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

## The code

```{r}
#| eval: false
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```
:::

I can also specify quantiles by cut points such as 2.5% and 97.5% tails to colour the ridgeline plot as shown in the figure/code chunk below.

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```

## **Visualising Distribution with Raincloud Plot**

### **Plotting a Half Eye graph**

First, we will plot a Half-Eye graph by using [`stat_halfeye()`](https://mjskay.github.io/ggdist/reference/stat_halfeye.html) of **ggdist** package.

This produces a Half Eye visualization, which is contains a half-density and a slab-interval.

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

### **Adding the boxplot with `geom_boxplot()`**

Next, we will add the second geometry layer using [`geom_boxplot()`](https://r4va.netlify.app/chap09) of ggplot2. This produces a narrow boxplot. We reduce the width and adjust the opacity.

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

### **Adding the Dot Plots with `stat_dots()`**

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

### **Finishing touch**

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

# 2 **Visual Statistical Analysis**

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

### **One-sample test: *gghistostats()* method**

In the code chunk below, [*gghistostats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/gghistostats.html) is used to to build an visual of one-sample test on English scores.

```{r}

set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

### **How to interpret Bayes Factor**

### **Two-sample mean test: *ggbetweenstats()***

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

###  **Oneway ANOVA Test: *ggbetweenstats()* method**

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### **Significant Test of Correlation: *ggscatterstats()***

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

### **Significant Test of Association (Depedence) : *ggbarstats()* methods**

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

```{r}
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

## **Visualising Models**

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

# 3 **Visualising Uncertainty**

```{r}
#| eval: false
devtools::install_github("wilkelab/ungeviz",force = TRUE)
```

```{r}
pacman::p_load(ungeviz, plotly, crosstalk,
               DT, ggdist, ggridges,
               colorspace, gganimate, tidyverse)
```

## **Visualizing the uncertainty of point estimates: ggplot2 methods**

```{r}

my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
```

Next, the code chunk below will be used to display *my_sum* tibble data frame in an html table format.

```{r}

knitr::kable(head(my_sum), format = 'html')
```

### **Plotting standard error bars of point estimates**

Now I will plot the standard error bars of mean maths score by race as shown below.

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by rac")
```

### **Plotting confidence interval of point estimates**

```{r}

ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")
```

### **Visualizing the uncertainty of point estimates with interactive error bars**

```{r}

shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))
```

## **Visualising Uncertainty: ggdist package**

### **Visualizing the uncertainty of point estimates: ggdist methods**

```{r}

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

```{r}

exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")
```

### **Visualizing the uncertainty of point estimates: ggdist methods**

```{r}

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval(
    show.legend = FALSE) +   
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

### **Visualizing the uncertainty of point estimates: ggdist methods**

```{r}

exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

## **Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)**

```{r}
#| eval: false
devtools::install_github("wilkelab/ungeviz")
```

```{r}
library(ungeviz)
```

```{r}

ggplot(data = exam, 
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, width = 0.05), 
    size = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  # `.draw` is a generated column indicating the sample draw
  transition_states(.draw, 1, 3)
```

## **Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)**

```{r}

ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

# 4 **Funnel Plots for Fair Comparisons**
